---
title: "Description of corpus"
author: "Bodo"
date: "16/11/2020"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description of the corpus

This script merely characterizes the corpus. It works as a brief reference to get an overview of how many birds, words etc. there are:

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(tidytext) # for unnest_tokens()
```

Get all file names:

```{r message = FALSE, warning = FALSE}
birds <- read_csv('../data/birds_master_file.csv')
```

# Summary information

How many entries altogether?

```{r}
nrow(birds)
```

How many per field guide?

```{r}
birds |> count(fieldguide)
```

How many bird species?

```{r}
birds |> 
  count(latin_species, sort = TRUE)
```

How many per species?

```{r}
birds |> 
  count(latin_species, sort = TRUE) |> 
  summarize(mean_entries = mean(n))
```

Approximately 3 per species...

# Word count

Let's do word counts without onomatopoeia and Warblish:

```{r}
birds <- birds |> 
  mutate(voice_no_onom = str_replace_all(voice, '\\".+?\\"', ''),
         voice_no_onom = str_replace_all(voice_no_onom, '\\*.+?\\*', ''))
```

Count the tokens, excluding onomatopoeia and warblish:

```{r}
birds |> 
  unnest_tokens(word,
                voice_no_onom) |> 
  nrow()
```

138,338 tokens. Note that this includes genitive 's as separate word (I guess defensible since it is a particle/clitic), and excludes onomatopoeia and warblish.

Count the tokens, including onomatopoeia and warblish:

```{r}
birds |> 
  unnest_tokens(word,
                voice) |> 
  nrow()
```

159,630 words then.


