---
title: "Bird talk analysis: metadata / typographic highlighting"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load data

Part of the larger linguistic analysis of the language of bird vocalizations.

In this script, we'll merely look at the metadata, which includes information about authors etc., and whether and how onomatopoeias are typographically highlighted.

Load libraries:

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
```

Load the metadata:

```{r message = FALSE, warning = FALSE}
meta <- read_csv('../data/fieldguides_metadata.csv')
```

# Descriptive stats

Range of years across field guides:

```{r}
range(meta$current_edition)
```

How many authors?

```{r}
meta %>% 
  count(author, sort = TRUE)
```

17 distinct author teams.

Check typographic highlighting:

```{r}
# Plot basics:

typo_p <- meta %>% 
  count(onomatopoeia_highlighting, sort = TRUE) %>% 
  bind_rows(tibble(onomatopoeia_highlighting = 'no highlighting', n = 0)) %>% 
  ggplot(aes(x = reorder(onomatopoeia_highlighting, n), y = n)) +
  geom_col(fill = 'steelblue', col = 'black', width = 0.6)

# Axes and scales:

typo_p <- typo_p +
  xlab(NULL) +
  ylab('Number of field guides') +
  coord_cartesian(ylim = c(0, 12)) +
  scale_y_continuous(expand = c(0, 0),
                     breaks = seq(0, 12, 2))

# Cosmetics:

typo_p <- typo_p +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        axis.title.y = element_text(face = 'bold',
                                    margin = margin(r = 10)))

# Show and save:

typo_p
ggsave(plot = typo_p,
       filename = '../figures/typographic_highlighting.pdf',
       width = 5, height = 3)
```

This completes this analysis.
